---
title: "Reproducible Research: Peer Assessment 1"
author : Mithun IM
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r,echo=TRUE}

data <- read.csv('activity.csv')
data_filled <- data 
data_step <- data[,1]
data_interval <- data[,3]
datastep_no_missing <- data_step[!is.na(data_step)==TRUE]
datainterval_no_missing <- data_interval[!is.na(data_step)==TRUE]

```

## What is mean total number of steps taken per day?

```{r,echo=TRUE}
total_steps_per_day <- tapply(data$steps, data$date, sum, na.rm=TRUE)
total_steps_per_day
mean(total_steps_per_day)
median(total_steps_per_day)
hist(total_steps_per_day,breaks = 15) 
```

## What is the average daily activity pattern?

```{r, echo = TRUE}
avg_steps_per_interval <- tapply(data$steps, data$interval, mean, na.rm=TRUE) # avg_steps_per_interval
data_interval_unique <- unique(data$interval)
plot(data_interval_unique,avg_steps_per_interval,type = "l")
```

```{r, echo = TRUE}
max_avg <- avg_steps_per_interval[avg_steps_per_interval==max(avg_steps_per_interval)]
max_interval <- which(avg_steps_per_interval==max(avg_steps_per_interval))
```

The maximum number of steps taken on a day was `r max_avg` on `r max_interval`th 5 minute interval.

## Imputing missing values

```{r, echo=TRUE}
a <- sum(is.na(data[,1])) # Total number of missing days
a
```
There are 288 5 minute intervals in a day. Data is missing on 8 days completely (2304/288=8).

To replace the missing values, mean of the 5 minute intervals is used.
```{r,echo = TRUE}
data_filled$steps[which(is.na(data$steps))] <- avg_steps_per_interval
```
New dataset with filled values is data_filled. 
```{r,echo = TRUE}
steps_per_day_filled <- tapply(data_filled$steps, data_filled$date, sum, na.rm=TRUE)
mean(steps_per_day_filled)
median(steps_per_day_filled)
hist(steps_per_day_filled,breaks = 15) 
``` 

The mean and median has coincided now. Both Mean and Median increased. Total steps per day has also increased. 

## Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}

data_filled_week <- weekdays((as.POSIXct(data_filled[,2])))
weekend_weekday <- factor(data_filled_week %in% c("Saturday","Sunday"), 
               labels=c("weekday","weekend"))
week_total_steps <- tapply(data_filled$steps,weekend_weekday, sum)
weekday_avg <- week_total_steps[1]/5
weekday_avg
weekend_avg <- week_total_steps[2]/2
weekend_avg

data_weekend <- data_filled[weekdays(as.POSIXct(data_filled[,2])) %in% c("Sunday","Saturday"),]
data_weekday <- data_filled[weekdays(as.POSIXct(data_filled[,2])) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),]
avg_steps_weekday <- tapply(data_weekday$steps, data_weekday$interval, mean, na.rm=TRUE) 
avg_steps_weekend <- tapply(data_weekend$steps, data_weekend$interval, mean, na.rm=TRUE) #
plot(avg_steps_weekday,type = "l")
lines(avg_steps_weekend,type = "l",col = "red",xaxt="n")
legend("topright",c("weekday","weekend"),bty = "n",lty = 1)
```

Number of steps taken during weekends are more compared to weekdays.
```